Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-Occurrence across \nOntario’s Far North

### 2023-08-21 analysis/paper/paper.qmd
updates

if (!file.exists("analysis/data/derived_data/annotations.rds") |
    !file.exists("analysis/data/derived_data/annotations_wide.rds")){
  source("scripts/1_preprocess_annotations.R")
} else {
  anns <- readRDS("analysis/data/derived_data/annotations.rds")
  anns_wide <- readRDS("analysis/data/derived_data/annotations_wide.rds")
}

if (!file.exists("analysis/data/derived_data/detections.rds")){
  source("scripts/2_preprocess_detections.R")
} else {
  dets <- readRDS("analysis/data/derived_data/detections.rds")
}

@@

>>>>

if (!file.exists("analysis/data/derived_data/annotations.rds") |
    !file.exists("analysis/data/derived_data/annotations_wide.rds")){
  # source("scripts/1_preprocess_annotations.R")
} else {
  # anns <- readRDS("../analysis/data/derived_data/annotations.rds")
  # anns_wide <- readRDS("analysis/data/derived_data/annotations_wide.rds")
  anns_wide <- readRDS("analysis/data/derived_data/annotations_wide_treated.rds")
}

if (!file.exists("analysis/data/derived_data/detections.rds")){
  # source("scripts/2_preprocess_detections.R")
} else {
  dets <- readRDS("analysis/data/derived_data/detections.rds")
}

@@

```{r}
# library(forcats)
# anns_wide_animals_joined_mut %>% 
#   ggplot(aes(x=date_time, y=species)) + geom_point()
# anns_wide_animals_joined_mut %>% 
#   dplyr::filter(species == "Fisher") %>% View()

camera_days <- exif %>% dplyr::group_by(maker_notes_user_label) %>% 
  dplyr::summarise(camera_first_time=min(exif_date_time_original)) %>% 
  dplyr::mutate(date_time = lubridate::ymd_hms(camera_first_time,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_")) %>% 
  dplyr::select(camera_id = maker_notes_user_label, 
                camera_first_day = date_time)

camera_regions <- cams_sf %>% sf::st_join(zones) %>% 
  sf::st_drop_geometry() %>% dplyr::select(camera_id, ECOZONE_NA)

bp <- anns_wide_animals_joined_mut %>%

  dplyr::filter(species != "Fisher") %>% 
  dplyr::left_join(camera_days, by = "camera_id") %>%

  dplyr::select(species, camera_id, 
                date_time, camera_first_day, ymd) %>%

  dplyr::group_by(species, camera_id) %>% 
  dplyr::summarise(first_day = min(date_time),
                   camera_first_day = min(camera_first_day),
                   camera_events = length(unique(ymd))) %>% 
  dplyr::ungroup() %>%

  dplyr::left_join(camera_regions, by = "camera_id") %>%

  dplyr::group_by(species) %>% 
  dplyr::mutate(total_events = sum(camera_events),
                n_cameras = dplyr::n()) %>% 
  dplyr::ungroup() %>%

  # dplyr::filter(n_cameras > 4) %>%
  dplyr::filter(n_cameras > 3) %>%

  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
                ECOZONE_NA = as.factor(ECOZONE_NA)) %>%

  ggplot(aes(x = diff_time, 
             y = forcats::fct_reorder(species, total_events),
             color = ECOZONE_NA)) +
  geom_boxplot() +

  scale_x_continuous()+
  labs(x = "Detection time (in days)",
       y = "Species",
       color = "Ecozone") + 
  theme_bw()
ggsave(plot = bp, filename = "analysis/figures/bp.png", width = 10, height = 15)

pts <- anns_wide_animals_joined_mut %>%

  dplyr::filter(species != "Fisher") %>% 
  dplyr::left_join(camera_days, by = "camera_id") %>%

  dplyr::select(species, camera_id, 
                date_time, camera_first_day, ymd) %>%

  dplyr::group_by(species, camera_id) %>% 
  dplyr::summarise(first_day = min(date_time),
                   camera_first_day = min(camera_first_day),
                   camera_events = length(unique(ymd))) %>% 
  dplyr::ungroup() %>%

  dplyr::left_join(camera_regions, by = "camera_id") %>%

  dplyr::group_by(species) %>% 
  dplyr::mutate(total_events = sum(camera_events),
                n_cameras = dplyr::n()) %>% 
  dplyr::ungroup() %>%

  # dplyr::filter(n_cameras > 4) %>%
  dplyr::filter(n_cameras > 3) %>%

  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
                ECOZONE_NA = as.factor(ECOZONE_NA)) %>%

  ggplot(aes(x = diff_time, 
             y = forcats::fct_reorder(species, total_events),
             # size = camera_events, 
             color = ECOZONE_NA)) +

  geom_point(aes(x = diff_time, 
                  y = forcats::fct_reorder(species, total_events),
                  size = camera_events, 
                  color = ECOZONE_NA,
                  group = ECOZONE_NA)) +

  scale_x_continuous()+
  labs(x = "Detection time (in days)",
       y = "Species",
       color = "Ecozone",
       size = "Day counts") + 
  theme_bw()
ggsave(plot = pts, filename = "analysis/figures/pts.png", width = 10, height = 15)
pts
```

```{r}
for_vic <- anns_wide_animals_joined_mut %>% 
  dplyr::select(id,species,date_time, camera_id) %>% 
  dplyr::left_join(cams_sf %>% dplyr::select(camera_id, plot_id) %>% sf::st_drop_geometry(),
                   by = "camera_id")
tot_camera <- for_vic$camera_id %>% unique() %>% length()
hit_rates <- for_vic %>% dplyr::select(species, camera_id) %>% 
  dplyr::group_by(species) %>% dplyr::summarise(n_cameras=length(unique(camera_id))) %>% 
  dplyr::ungroup() %>% dplyr::mutate(rate = n_cameras/tot_camera*100) %>% 
  dplyr::arrange(dplyr::desc(rate))
# det_history_two_weeks <-

readr::write_csv(for_vic, "../data/derived_data/observations_by_camera_2023Aug15.csv")
readr::write_csv(hit_rates, "../data/derived_data/hit_rates_by_species_2023Aug15.csv")
```
